---
title: "NFL Analysis"
# description: |
#   A new article created using the Distill format.
# author:
#   - name: Nora Jones 
#     url: https://example.com/norajones
#     affiliation: Spacely Sprockets
#     affiliation_url: https://example.com/spacelysprokets
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(nflfastR)
library(tidyverse)

pbp <- nflfastR::load_pbp(2022) %>% 
  filter(season_type == "REG") %>% 
  filter(!is.na(posteam) & (rush == 1 | pass == 1))

offense <- pbp %>%
  dplyr::group_by(team = posteam) %>%
  dplyr::summarise(off_epa = mean(epa, na.rm = TRUE))

defense <- pbp %>%
  dplyr::group_by(team = defteam) %>%
  dplyr::summarise(def_epa = mean(epa, na.rm = TRUE))
```

# Introduction

My favorite team is the New England Patriots. The Patriots had a bit of a down year 
in the 2022 season. This analysis is intended to dive into some of the data to get 
to the bottom of why the team struggled.


# Offense

Many point to the offense as the main reason for the Patriots struggles, and rightly 
so as they came into the season with questions about their offensive coaching staff 
and how that would impact their 2nd year quarterback, Mac Jones.  

```{r}
offense %>%
  dplyr::inner_join(defense, by = "team") %>%
  ggplot2::ggplot(aes(x = off_epa, y = def_epa)) +
  ggplot2::geom_abline(slope = -1.5, intercept = c(.4, .3, .2, .1, 0, -.1, -.2, -.3), alpha = .2) +
  nflplotR::geom_mean_lines(aes(h_var = off_epa, v_var = def_epa)) +
  nflplotR::geom_nfl_logos(aes(team_abbr = team), width = 0.07, alpha = 0.7) +
  ggplot2::labs(
    x = "Offense EPA/play",
    y = "Defense EPA/play",
    caption = "Data: @nflfastR",
    title = "2022 NFL Offensive and Defensive EPA per Play"
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    plot.title = ggplot2::element_text(size = 12, hjust = 0.5, face = "bold")
  ) +
  ggplot2::scale_y_reverse()
```

<br>

As we can see here, the defense was near top of the league in EPA, while the offense 
ranked below the league average.




